---
title: "Ec140 - Core Concepts from Statistics"
author: "Fernando Hoces la Guardia"
date: "06/22/2022"
output: 
  xaringan::moon_reader:
    footer: "These slides available at https://fhoces.github.io/econ140summer2022/"
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---


<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 1em 1em 1em;
}
</style>


```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, viridis, knitr, extrafont, tidyverse, magrittr, latex2exp, parallel, Ecdat, wooldridge, dslabs, ggforce)
# Define colors
red_pink <- "#e64173"
met_slate <- "#23373b" # metropolis font color
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  warning = F,
  message = F
)  
theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 14),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
theme_empty <- theme_bw() + theme(
  line = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
```

# Housekeeping 

- Dates for chapter summaries are now on the syllabus. 

- Any troubles accessing the book?

- Any first thoughs on PS 1?

- Font size for people in the back ok?

---
# Concepts to Review Today
.font140[
- Random Variables 

- Probability and Probability Distribution/Density Functions

]

---
# Random Variables 1/4

Imagine you are meeting with a friend later today and let’s assume of the three following situations can happen: 
- You have the most wonderful time that you could have had today with your friend.
- It was ok, but you were kind of expecting to have a better time today. 
- You had a really bad time, your friend did not stop talking about herself and never checked on you.

--

Roll one Die: 

- If you get a 1 or a 2, the first situation happen
- If you get a 3 or a 4, the second situation took place
- 5 or 6 corresponds to the last situation



---
# Random Variables 2/4
.pull-left[
Each of these situations is what is called an **event** in statistics. Two key characteristics of events are
  - They must be mutually elusive
  - The collection of all events (called the event space) must contain all the posible outcomes. 

This events look a bit hard to keep track of...
]
--




.pull-right[

Let's assign a number to each event: 

```{r, echo=FALSE, include=TRUE, results="asis"}

# Generate data
df <- data.frame(
  event = c(
    "You have the most wonderful...", 
    "It was ok, but...", 
    "You had a really bad time, ..."
  ),
  x = c(1, 2, 3) # draw integers
)

colnames(df) <- c("\\(Event\\)", "\\(X\\)")

df %>% DT::datatable(caption = "Numbering Events", options = list(dom = "t", ordering = FALSE, columnDefs = list(list(className = 'dt-center', targets = "_all"))), rownames = FALSE)
```
]

---
# Random Variables 3/4

.pull-left[

- This assignment from event to numerical values is what a **random variable** does. 

- For this case, where there is a discrete number events, we called this a  **discrete random variable**

- You might be used to thinking about variables as collections of numbers in a line (or multiple lines). 


]


.pull-right[

```{r, echo=FALSE, include=TRUE, results="asis"}

# Generate data
df <- data.frame(
  event = c(
    "You have the most wonderful...", 
    "It was ok, but...", 
    "You had a really bad time, ..."
  ),
  x = c(1, 2, 3) # draw integers
)

colnames(df) <- c("\\(Event\\)", "\\(X\\)")

df %>% DT::datatable(caption = "Numbering Events", options = list(dom = "t", ordering = FALSE, columnDefs = list(list(className = 'dt-center', targets = "_all"))), rownames = FALSE)
```

- A random variable is similar in the sense that it is a collection of numbers, but it is different in the sense that each number represents an event. 

]


---
# Random Variables 4/4


- Another potential source of confusion is that the idea of connecting something with other numbers is similar of the role of a function, not variables. We need random variables first to move from events into numbers.

--

- Random variables can also be continued, but for this we cannot talk of an Event-value mapping, instead we combine the event and value in the same object (the number of the random var.). 
    - Example: income. In the continuous case, the idea of an event is not so meaningful (what is the likelihood of an income of 33,593.4355?) 
    - Hence, when working with continuous random variables we will refer to ranges. 

---

# Seeing Theory

  - Let's go to [Seeing Theory](https://seeing-theory.brown.edu/probability-distributions/index.html#section1) and play a bit with a discrete random variable
      - For example: think of that event space as a field where you are throwing a ball
--

Now we have all we need to define a random variable. 

- **Random Variable** is a numerical summary of a random outcome. 

--

- We now know what a random variable is. Notice that we have almost not talk about how likely or unlikely each of these values are, we have not talked about its *proportions in the long run*. 


---
# Probability Distribution Functions  1/4

- "Proportions in the long run", is a bit of a mouthful. Let’s put a name to it, let’s call it **probabilities**. 

- Now we need a mapping from an event (“Most wonderful time that you could have had today with your friend”, or 1) into a probability. 

---
# Probability Distribution Functions  2/4


- This mapping will be described by a function $P(X=x)$ that represents the probability that a random variable $X$, takes the specific value of $x$ (e.g., 1, or "Most wonderful ...").

```{r, echo=FALSE, include=TRUE, results="asis"}

# Generate data
df <- data.frame(
  event = c(
    "You have the most wonderful...", 
    "It was ok, but...", 
    "You had a really bad time, ..."
  ),
  x = c(1, 2, 3) , 
  px = c("1/3", "1/3", "1/3")# draw integers
)

colnames(df) <- c("\\(Event\\)", "\\(X\\)", "\\(P(X = x)\\)")

df %>% DT::datatable(caption = "Events, Random Variables, and Probabilities", options = list(dom = "t", ordering = FALSE, columnDefs = list(list(className = 'dt-center', targets = "_all"))), rownames = FALSE)
```

---
# Probability Distribution Functions  3/4

 - The **probability distribution function** of a discrete random variable is the list of all possible values of a variable and the probability that each value will occur. 
 
 - Notation. Sometimes you will see it express as $P(X=x)$ or as $f(x)$. The former applies only to discrete PDFs, the latter can be used for both continous or discrete random variables. 
 
 - A closely related concept describes the acumulated probability of a collection of events. How would you write the probability that people in this room had "At least an ok time with their friend"
 
 
---
# Probability Distribution Functions  4/4

- This Cumulative Distribution Function describes the probability that a random variables is less or equal than a particular value. $P(X\leq x)$

- Lets go to [Seeing Theory](https://seeing-theory.brown.edu/probability-distributions/index.html#section1) again and explore some PDFs and CDFs of discrete random variables. 
    - Bernoulli: success and failure, 0 and 1. Show how p characterizes the entire distributions. Review ranges and values of PDF and CDF. 
    - Binomial as example of combining RVs.




---
# Activity 1 


 - The mythical [Stat 110 team](https://projects.iq.harvard.edu/stat110/home) shows us how the concepts of Discrete Random Variables and Probability Distributions can be used in practice. 
 - Let's watch [this video](https://youtu.be/ZoIPuTIPviY). Then get together in groups of 3 and discuss:
    - What is the event space?
    - What is the random variable?
    - What is the probability distribution that one person is cured?
    - How does the Binomial(10, 0.6) relates to that previous distribution?



---
# Probability Density Functions 1/2

- For the case of continuous random variables, lets think of what is the probability of a specific event (e.g., income 34,680.0003). 

--

- For continued random variables we cannot talk about a (specific) probability distribution, we use instead the concepts of density (as in how much is there in any given range). 

- The probability that a continues random variable lies between any given range (of two points) is defined as the area under the **probability density function**  of a continuous random variable. [This is a very abstract concept] 

- The density function is typically denoted by $f(x)$. 


---
# Probability Density Functions 2/2


- Note that any specific value of $f(x)$ should not be interpreted as a probability. 

- The area under the the probability density function, or density, is the cumulative distribution function between two points: 

$$ 
\begin{equation}
P(x_{1} \leq X \leq x_{2}) = \text{area betwen } f(x_{1}) \text{ and } f(x_{2})\\
\end{equation}
$$

---

# Seeing Theory 2

- Lets go to [Seeing Theory](https://seeing-theory.brown.edu/probability-distributions/index.html#section1) again and explore some PDFs and CDFs of continous random variables. 
    - Uniform: great to represent ignorance.
    - Normal to show how just two parameters can characterize and entire prob distribution. 
    - Exponential to show that weird shapes can happen. [Check this video](https://youtu.be/XXjlR2OK1kM) for a great demonstration of how this type of distributions appear in real life.
    - Beta to show that one distribution can describe many phenomena. Talks about range.

---

# Activity 2
 - Let's watch [another Stat 110's video](https://youtu.be/UVQs9zikfe0). Then get together in groups of 3 and discuss:
    - What is the type of random variables are represented by Norma and Randy?
    - What is Norma's take on how to address the paradox describe by ancient philosophers?
    - The concept of error tolerance presented in the video correspnds to which concept discuss in previous slides?
    - The last 2 mins on prediction are outside the scope of our course. 





---
# Densities of Continuous Random Variables

.pull-left[
## Uniform Distribution

The probability density function of a variable uniformly distributed between 0 and 2 is

$$
f(x) =
\begin{cases}
  \frac{1}{2} & \text{if } 0 \leq x \leq 2 \\
  0 & \text{if } x < 0 \text{ or } x>2
\end{cases}
$$
]

.pull-right[
```{R, unif, echo = F, dev = "svg", fig.height = 6}
x <- seq(-1, 3, 0.01)
y <- seq(1, 1.5, 0.01)
z <- seq(0, 2 ,0.01)

ggplot() +
  scale_x_continuous(limits = c(-1, 3), expand=c(0,0)) +
  scale_y_continuous(limits = c(0, 1), expand=c(0,0), breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1)) +
  geom_ribbon(aes(x = z, ymin = 0, ymax = 0.5), fill = "grey85", linetype = "blank") +
  xlab(expression(X)) + 
  ylab("Density") +
  theme_simple
```
]
---
# Densities of Continuous Random Variables

.pull-left[
## Uniform Distribution

By definition, the area under $f(x)$ is equal to 1.

The .hi[shaded area] illustrates the probability of the event $1 \leq X \leq 1.5$.

- $\mathop{\mathbb{P}}(1 \leq X \leq 1.5) = (1.5-1) \times0.5 = 0.25$.
]

.pull-right[
```{R, unif2, echo = F, dev = "svg", fig.height = 5}
x <- seq(-1, 3, 0.01)
y <- seq(1, 1.5, 0.01)
z <- seq(0, 2 ,0.01)

ggplot() +
  scale_x_continuous(limits = c(-1, 3), expand=c(0,0)) +
  scale_y_continuous(limits = c(0, 1), expand=c(0,0), breaks = c(0, 0.5, 1), labels = c(0, 0.5, 1)) +
  geom_ribbon(aes(x = z, ymin = 0, ymax = 0.5), fill = "grey85", linetype = "blank") +
  geom_ribbon(aes(x = y, ymin = 0, ymax = 0.5), fill = red_pink, linetype = "blank") +
  geom_vline(xintercept = 1, size = 0.35, linetype = "dashed", color = met_slate) +
  geom_vline(xintercept = 1.5, size = 0.35, linetype = "dashed", color = met_slate) +
  xlab(expression(X)) + 
  ylab("Density") +
  theme_simple
```
]
---
# Densities of Continuous Random Variables

.pull-left[
## Normal Distribution

.hi-purple[The "bell curve."]

- Symmetric: mean and median occur at the same point (_i.e._, no skew).

- Low-probability events in tails; high-probability events near center.

]

.pull-right[

```{R, norm, echo = F, dev = "svg", fig.height = 5}
df <- tibble(x = seq(-4, 4, 0.01), y = dnorm(x)) %>%
  rbind(., tibble(x = seq(4, -4, -0.01), y = 0))
ggplot() +
  scale_x_continuous(limits = c(-4, 4), expand=c(0,0)) +
  scale_y_continuous(limits = c(0, 0.5), expand=c(0,0), breaks = c(0, 0.5), labels = c(0, 0.5)) +
  geom_polygon(data = df, aes(x, y), fill = "grey85") +
  #geom_polygon(data = df %>% filter(x <= qnorm(0.5 + 0.475) & x >= qnorm(0.5 - 0.475)), aes(x, y), fill = red_pink) +
  #geom_vline(xintercept = qnorm(0.5 + 0.475), size = 0.35, linetype = "dashed", color = met_slate) +
  #geom_vline(xintercept = qnorm(0.5 - 0.475), size = 0.35, linetype = "dashed", color = met_slate) +
  theme_simple +
  xlab(expression(X)) + 
  ylab("Density")
```
]
---
# Continuous Random Variables

.pull-left[
## Normal Distribution

The .hi[shaded area] illustrates the probability of the event $-2 \leq X \leq 2$.

- "Find area under curve" .mono[=] use integral calculus (or, in practice, .mono[R]).

- $\mathop{\mathbb{P}}(-2 \leq X \leq 2) \approx 0.95$. 

]

.pull-right[
```{R, norm2, echo = F, dev = "svg", fig.height = 5}
df <- tibble(x = seq(-4, 4, 0.01), y = dnorm(x)) %>%
  rbind(., tibble(x = seq(4, -4, -0.01), y = 0))
ggplot() +
  scale_x_continuous(limits = c(-4, 4), expand=c(0,0)) +
  scale_y_continuous(limits = c(0, 0.5), expand=c(0,0), breaks = c(0, 0.5), labels = c(0, 0.5)) +
  geom_polygon(data = df, aes(x, y), fill = "grey85") +
  geom_polygon(data = df %>% filter(x <= qnorm(0.5 + 0.475) & x >= qnorm(0.5 - 0.475)), aes(x, y), fill = red_pink) +
  geom_vline(xintercept = qnorm(0.5 + 0.475), size = 0.35, linetype = "dashed", color = met_slate) +
  geom_vline(xintercept = qnorm(0.5 - 0.475), size = 0.35, linetype = "dashed", color = met_slate) +
  theme_simple +
  xlab(expression(X)) + 
  ylab("Density")
```

]
---
# Acknowledgments

- I started this slides on the basis of [Kyle Raze's slides](https://raw.githack.com/kyleraze/EC320_Econometrics/master/Lectures/02-Statistics_Review/02-Statistics_Review.html#34) from University of Oregon.  
- [Seeing Theory](https://seeing-theory.brown.edu/index.html). 
- [Stat 110](https://projects.iq.harvard.edu/stat110/home). 
- Stock and Watson 3e. Chapter 2, P61-65. 
- [Numberphile!](https://www.numberphile.com)


```{r gen_pdf, include = FALSE, cache = FALSE, eval = FALSE}
pagedown::chrome_print("02_stats_concepts.html", output = "02_stats_concepts.pdf")
```