---
title: "Ec140 - Midterm 1 Review"
author: "Fernando Hoces la Guardia"
date: "07/11/2022"
output: 
  xaringan::moon_reader:
    footer: "These slides available at https://fhoces.github.io/econ140summer2022/"
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 1em 1em 1em;
}
</style>

```{r , include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, viridis, knitr, extrafont, tidyverse, magrittr, latex2exp, 
       parallel, Ecdat, wooldridge, dslabs, ggforce, emo, png, grid, pander, 
       countdown, emoGG, haven, broom)


options(htmltools.dir.version = FALSE)

# define vars
om = par("mar")
lowtop = c(om[1],om[2],0.1,om[4])

overwrite = FALSE


# Define colors
red_pink <- "#e64173"
met_slate <- "#23373b" # metropolis font color
# Notes directory
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  message = FALSE,
  warning = FALSE,
  dev = "svg",
  cache = TRUE
  #fig.width = 11,
  #fig.height = 5
)

theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 14),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
theme_empty <- theme_bw() + theme(
  line = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)

# countdown style
countdown(
  color_border              = "#d90502",
  color_text                = "black",
  color_running_background  = "#d90502",
  color_running_text        = "white",
  color_finished_background = "white",
  color_finished_text       = "#d90502",
  color_finished_border     = "#d90502"
)



```



# Midterm Review
- Go over the exam. 

---

# Midterm Takeaways 1/2

My expectations: 

- If you: 
  -  (1) actively engage with class material, **and** 
  -  (2) read the book, **and**
  -  (3) did section + practice questions, **and** 
  -  (4) check with classmates and/or us in OH, 
  **then**
  - you should have done well in the midterm. 

- Using some of the course language: conditional on doing the work, your expected permanence should be high. 
---

# Midterm Takeaways 2/2

- If you did (1) - (4) and did not do well, probably it was a bad day, please come to OH and let's work on a plan so this doesn't happen again. 

- If you did not do either (1), (2), (3), or (4) and did not do well, do them and I am confident that you will do much better in Midterm 2 and Exam.

- Final comment: I hope that it's clear now that "consumer approach" $\neq$ "easy course". It means: if you do the work, you will leave this course with a strong foundation to read and use causal evidence and contribute to policy discussions (and also with strong foundation for further studies in econometrics). 




```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
pagedown::chrome_print("10_midterm_review.html", output = "10_midterm_review.pdf")
```