---
title: "Last Thoughts"
author: "Fernando Hoces la Guardia"
date: "08/10/2022"
output: 
  xaringan::moon_reader:
    footer: "These slides available at https://fhoces.github.io/econ140summer2022/"
    css: [default, metropolis, metropolis-fonts, "my-css.css"] 
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---
count:false

<style type="text/css">
.remark-slide-content {
    font-size: 30px;
    padding: 1em 1em 1em 1em;
}
</style>

```{r , include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, viridis, knitr, extrafont, tidyverse, magrittr, latex2exp, 
       parallel, Ecdat, wooldridge, dslabs, ggforce, emo, png, grid, pander, 
       countdown, emoGG, haven, broom, ggridges, gridExtra, kableExtra, snakecase, 
       janitor, data.table, lubridate, lfe, here, gapminder, AER, stargazer, margins, 
       furrr)


# Define pink color
red_pink <- "#e64173"
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#3b3b9a"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"
# Dark slate grey: #314f4f

options(htmltools.dir.version = FALSE)

# define vars
om = par("mar")
lowtop = c(om[1],om[2],0.1,om[4])

overwrite = FALSE


# Define colors
red_pink <- "#e64173"
met_slate <- "#23373b" # metropolis font color
# Notes directory
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  message = FALSE,
  warning = FALSE,
  dev = "svg",
  cache = FALSE
  #fig.width = 11,
  #fig.height = 5
)

theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 14),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
theme_empty <- theme_bw() + theme(
  line = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  plot.margin = structure(c(0, 0, -0.5, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_simple <- theme_bw() + theme(
  line = element_blank(),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text.x = element_text(size = 18, family = "STIXGeneral"),
  axis.text.y = element_blank(),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  # plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_axes_math <- theme_void() + theme(
  text = element_text(family = "MathJax_Math"),
  axis.title = element_text(size = 22),
  axis.title.x = element_text(hjust = .95, margin = margin(0.15, 0, 0, 0, unit = "lines")),
  axis.title.y = element_text(vjust = .95, margin = margin(0, 0.15, 0, 0, unit = "lines")),
  axis.line = element_line(
    color = "grey70",
    size = 0.25,
    arrow = arrow(angle = 30, length = unit(0.15, "inches")
  )),
  plot.margin = structure(c(1, 0, 1, 0), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_axes_serif <- theme_void() + theme(
  text = element_text(family = "MathJax_Main"),
  axis.title = element_text(size = 22),
  axis.title.x = element_text(hjust = .95, margin = margin(0.15, 0, 0, 0, unit = "lines")),
  axis.title.y = element_text(vjust = .95, margin = margin(0, 0.15, 0, 0, unit = "lines")),
  axis.line = element_line(
    color = "grey70",
    size = 0.25,
    arrow = arrow(angle = 30, length = unit(0.15, "inches")
  )),
  plot.margin = structure(c(1, 0, 1, 0), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_axes <- theme_void() + theme(
  text = element_text(family = "Fira Sans Book"),
  axis.title = element_text(size = 18),
  axis.title.x = element_text(hjust = .95, margin = margin(0.15, 0, 0, 0, unit = "lines")),
  axis.title.y = element_text(vjust = .95, margin = margin(0, 0.15, 0, 0, unit = "lines")),
  axis.line = element_line(
    color = grey_light,
    size = 0.25,
    arrow = arrow(angle = 30, length = unit(0.15, "inches")
  )),
  plot.margin = structure(c(1, 0, 1, 0), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_set(theme_gray(base_size = 20))
# Column names for regression results
reg_columns <- c("Term", "Est.", "S.E.", "t stat.", "p-Value")
# Function for formatting p values
format_pvi <- function(pv) {
  return(ifelse(
    pv < 0.0001,
    "<0.0001",
    round(pv, 4) %>% format(scientific = F)
  ))
}
format_pv <- function(pvs) lapply(X = pvs, FUN = format_pvi) %>% unlist()
# Tidy regression results table
tidy_table <- function(x, terms, highlight_row = 1, highlight_color = "black", highlight_bold = T, digits = c(NA, 3, 3, 2, 5), title = NULL) {
  x %>%
    tidy() %>%
    select(1:5) %>%
    mutate(
      term = terms,
      p.value = p.value %>% format_pv()
    ) %>%
    kable(
      col.names = reg_columns,
      escape = F,
      digits = digits,
      caption = title
    ) %>%
    kable_styling(font_size = 20) %>%
    row_spec(1:nrow(tidy(x)), background = "white") %>%
    row_spec(highlight_row, bold = highlight_bold, color = highlight_color)
}


```



```{css, echo = F, eval = T}
@media print {
  .has-continuation {
    display: block !important;
  }
}
```

# Key concepts
.font90[
.pull-left[
**Reinforced**


- Expectation/Mean  
- Standard deviation
- Law of large number
- Central limit theorem
- Conditional expectation
- Populations/Samples
- Hypothesis test/P-value

**New** 

- Selection Bias
- Potential outcomes

]

.pull-right[
- Independence
- Randomized Control Trials
- Regression as matching
- Regression as conditional expectation
- Regression as line fitting
- Regression anatomy
- Omitted variable bias
- Collinearity
- External validity
- Instrumental variables
- Regression discontinuity
- Difference in Difference
- Bad controls

]

]
---
count: true
background-image: url("Images/pre-cred-ebp.svg")
background-size: 70%
background-position: 50% 80%

# .font90[Credibility is Increasing in the Evidence-to-Policy Pipeline]
.center[
.font120[**Pre-Credibility** -- Cred. Revo. (1990s) -- Open Science+ (2010s) -->]
]
---
count: true
background-image: url("Images/pre-open-sci.svg")
background-size: 70%
background-position: 50% 80%

# What's wrong with the Evidence-to-Policy Pipeline?
.center[
.font120[Pre-Credibility -- **Cred. Revo. (1990s)** -- Open Science+ (2010s) -->]
]

.font80[
.pull-left[
Emphasis on empirically testing causal   
claims with clearly defined   
methods. See   
David Card’s    
[Nobel lecture](https://doi.org/10.1257/aer.112.6.1773).   
(Ec142)

<br>
Further from  
physics, closer   
to medicine   
([Chetty’s op ed. on NYT](https://www.nytimes.com/2013/10/21/opinion/yes-economics-is-a-science.html))
]
]

---
count: true
background-image: url("Images/pre-opa.svg")
background-size: 70%
background-position: 50% 80%

# What's wrong with the Evidence-to-Policy Pipeline?
.center[
.font120[Pre-Credibility -- Cred. Revo. (1990s) -- **Open Science+ (2010s)** -->]
]
.font80[
.pull-left[
Emphasis on empirically testing causal   
claims with clearly defined   
methods. See   
David Card’s    
[Nobel lecture](https://doi.org/10.1257/aer.112.6.1773).   
(Ec142)

<br>
Further from  
physics, closer   
to medicine   
([Chetty’s op ed. on NYT](https://www.nytimes.com/2013/10/21/opinion/yes-economics-is-a-science.html))
]
]
--
.font80[
.pull-right[
.right[
**[MY OPINION:]** Credibility revolution   
addresses one type of BS, but it  
completely disregards  
(and maybe reinforces)   
a second type of BS:   
the belief that  
we know how to     
use past research   
to inform   
current policy   
debates (aggregation   
and extrapolation of research)
]

]

]
---
# The Goal of EC140

<br><br>

.center[
.font130[
This course gave you the tools to think about causal evidence, to explore further and generate your own causal evidence. And, hopefully, to contribute and innovate to the space of connecting evidence with policy. 

]
]

---
# Last Warning

<br><br>

.center[
.font130[
Beware of Inaction!

]
]
---
# Acknowledgments (for Course Content)
.pull-left[
- Angrist and Pischke's Mastering Metrics
- Stock and Watson's Introduction to Econometrics
- Ed Rubin's multiple great courses
- Hoai-Luu Nguyen's Econoimate
- Nick Huntington-Klein multiple teaching resources
- Florian Oswald course on econometrics at Science Po
- XQCD
]

.pull-right[
- Joe Blitztein's Stat 110
- Seeing Theory
- Kyle Raze's course on econometrics
- Numberphile
- Matt Hollian's teaching resources for mastering metrics
- Eddie Woo's great explanations for statistics and probability
- Jeffrey Arnold's R companion for Matering Metrics
- Andrew Heiss courses on program evaluation
]

---
class: inverse, middle, center

.font140[
# Thank You and Good Luck!
]





```{r gen_pdf, include = FALSE, cache = FALSE, eval = TRUE}
pagedown::chrome_print("27_last_thoughts.html", output = "27_last_thoughts.pdf")
```